---
title: "Epicurves"
author: "Richard"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 6
vignette: >
  %\VignetteIndexEntry{Epicurves}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(ggplot2)
library(data.table)
library(fhiplot)
```

## Data

We generally start with a linelist dataset.

```{r}
set.seed(4)
dates <- sample(seq.Date(as.Date("2018-01-01"), as.Date("2018-02-08"), 1), 200, replace = T)
d <- data.frame(
  date = dates
)
print(d)
```

We now:

1. Convert this into a `data.table`
2. Aggregate it to daily data (this dataset doesn't contain any days with 0 cases)
3. Create a 'skeleton' dataset that contains all days from the first to last day
4. Merge the two datasets together (so you now have a dataset that contains days with 0 cases)
5. Fill in 'missing' N's with 0 (they are the dates that came from the skeleton dataset)

```{r}
# Convert to data.table
setDT(d)
# aggregate
d <- d[,
  .(
    N = .N
  ),
  keyby = .(
    date
  )
]
# aggregated daily dataset that does not contain days with 0 cases
print(d)

# create skeleton
skeleton <- data.table(
  date = seq.Date(min(d$date), max(d$date), 1)
)

# merge the two datasets together
d <- merge(d, skeleton, all = T)

# Fill in 'missing' Ns with 0
d[is.na(N), N := 0]

# Now you have a clean aggregated daily dataset that contains days with 0 cases!
print(d)
```

We can also create a weekly dataset:

```{r}
# create 3 new variables:

d[, isoyear := lubridate::isoyear(date)]
# we need 'formatC' to ensure that isoweek has padded 0s at the front
# so it is always 2 digits wide
d[, isoweek := formatC(lubridate::isoweek(date), width = 2, flag = 0)]

# paste isoyear and isoweek together
d[, isoyear_isoweek := glue::glue("{isoyear}-{isoweek}", isoyear = isoyear, isoweek = isoweek)]

# aggregate down to weekly level
w <- d[,
  .(
    N = sum(N)
  ),
  keyby = .(
    isoyear_isoweek
  )
]
print(w)
```

## Graphs

Daily epicurve

```{r}
q <- ggplot(d, aes(x = date, y = N))
q <- q + geom_col(fill = fhiplot::vals$cols$primary["B1"], width = 0.8)
q <- q + scale_x_date("Date")
q <- q + scale_y_continuous("Number of reported cases",
  breaks = scales::pretty_breaks(5),
  expand = c(0, 0)
)
q <- q + labs(title = "Epicurve from 2018-01-01 to 2018-02-20")
q <- q + labs(caption = "Data extracted on 2018-02-20")
q <- q + fhiplot::theme_fhi_lines()
q
```

Weekly epicurve

```{r}
q <- ggplot(w, aes(x = isoyear_isoweek, y = N))
q <- q + geom_col(fill = fhiplot::vals$cols$primary["B1"], width = 0.8)
q <- q + scale_x_discrete("Isoweek")
q <- q + scale_y_continuous("Number of reported cases",
  breaks = scales::pretty_breaks(5),
  expand = c(0, 0)
)
q <- q + labs(title = "Epicurve from 2018-01-01 to 2018-02-20")
q <- q + labs(caption = "Data extracted on 2018-02-20")
q <- q + fhiplot::theme_fhi_lines()
q
```

Weekly epicurve with vertical x-axis labels

```{r}
q <- ggplot(w, aes(x = isoyear_isoweek, y = N))
q <- q + geom_col(fill = fhiplot::vals$cols$primary["B1"], width = 0.8)
q <- q + scale_x_discrete("Isoweek")
q <- q + scale_y_continuous("Number of reported cases",
  breaks = scales::pretty_breaks(5),
  expand = c(0, 0)
)
q <- q + labs(title = "Epicurve from 2018-01-01 to 2018-02-20")
q <- q + labs(caption = "Data extracted on 2018-02-20")
q <- q + fhiplot::theme_fhi_lines()
q <- q + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
q
```
