---
title: "Palettes"
author: "Author"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 6
vignette: >
  %\VignetteIndexEntry{Palettes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(ggplot2)
library(fhiplot)
```

## Palette

```{r}
fhiplot::Display_All_Palettes()
```


## Points

We first try some scattergraphs with two different themes:

- White "classic" background
- Grey standard ggplot2 background

```{r}
ggplot(iris, aes(Sepal.Width, Sepal.Length, color = Species)) +
    geom_point(size = 4) +
    theme_classic()+
    fhiplot::scale_color_fhi(palette = "dis_primary")

ggplot(iris, aes(Sepal.Width, Sepal.Length, color = Species)) +
    geom_point(size = 4) +
    fhiplot::scale_color_fhi(palette = "dis_primary")

```

## Barcharts

We first try some barcharts with the `dis_primary` palette and two different themes:

- White "classic" background
- Grey standard ggplot2 background

```{r}
ggplot(diamonds, aes(x=color, fill = cut)) +
  theme_classic()+
  geom_bar() +
  fhiplot::scale_fill_fhi(palette = "dis_primary")

ggplot(diamonds, aes(x=color, fill = cut)) +
  geom_bar() +
  fhiplot::scale_fill_fhi(palette = "dis_primary")

```

We now try the `div_blue_red` palette:

```{r}
ggplot(diamonds, aes(x=color, fill = cut)) +
  theme_classic()+
  geom_bar() +
  fhiplot::scale_fill_fhi(palette = "div_blue_red", direction = -1)

ggplot(diamonds, aes(x=color, fill = cut)) +
  geom_bar() +
  fhiplot::scale_fill_fhi(palette = "div_blue_red", direction = -1)

```

We now try the `div_blue_red` palette with transparency set to `0.7`:

```{r}
ggplot(diamonds, aes(x=color, fill = cut)) +
  theme_classic()+
  geom_bar(alpha=0.7) +
  fhiplot::scale_fill_fhi(palette = "div_blue_red", direction = -1)

ggplot(diamonds, aes(x=color, fill = cut)) +
  geom_bar(alpha=0.7) +
  fhiplot::scale_fill_fhi(palette = "div_blue_red", direction = -1)
```

